---
title: "week10_lab paired t-test"
author: "Jian Zheng & ChatGPT"
date: '2023-04-06'
output: html_document
---

Check if tidyverse and ggplot2 are installed. If not, install the missing package(s).
```{r setup, include=FALSE}
if (!("tidyverse" %in% installed.packages()[, "Package"])) install.packages("tidyverse")
if (!("ggplot2" %in% installed.packages()[, "Package"])) install.packages("ggplot2")


library(tidyverse)
library(ggplot2)
```

# Introduction:
The paired t-test, also known as the dependent sample t-test, is a statistical method used to compare the means of two related samples. It is often used when there are two measurements taken on the same group, such as before and after an intervention. In this tutorial, we will go through the steps to perform a paired t-test in R

continue with the CO2 dataset, first let's have a look at the data
Half the plants of each type were chilled overnight before the experiment was conducted

# Prepare your data:
To perform a paired t-test, you will need two sets of related observations. In this tutorial, we will use a hypothetical dataset that compares the blood pressure of ten patients before and after a treatment.
```{r}
# Create sample data
patient_id <- c(1:10)
bp_before <- c(150, 145, 130, 120, 170, 160, 140, 135, 155, 165)
bp_after <- c(140, 130, 125, 115, 160, 155, 130, 120, 145, 155)

# Combine data into a data frame
bp_data <- data.frame(patient_id, bp_before, bp_after)
View(bp_data)
```

# Check the assumptions:
Before running the paired t-test, check the assumptions of normality and independence of observations. To test normality, we can use the Shapiro-Wilk test.

```{r}
# Test normality
shapiro_before <- shapiro.test(bp_data$bp_before)
shapiro_after <- shapiro.test(bp_data$bp_after)

print(shapiro_before)
print(shapiro_after)
```
If the p-values of both tests are greater than 0.05, we can assume normality.

# Perform the paired t-test:
Now that we have checked the assumptions, we can perform the paired t-test using the t.test() function in R

```{r}
# Perform paired t-test
paired_ttest <- t.test(bp_data$bp_before, bp_data$bp_after, paired = TRUE)
paired_ttest
```

We have saved the result of t-test into a list, now we can access each element by the name
```{r}
paired_ttest$p.value
paired_ttest$conf.int
```

# Interpret the results:
The output of the paired t-test will include the t-value, degrees of freedom, and p-value. If the p-value is less than your chosen significance level (e.g., 0.05), you can reject the null hypothesis and conclude that there is a significant difference between the means of the two related groups.

In our example, let's assume the p-value is less than 0.05. We can conclude that the treatment had a significant effect on blood pressure reduction.

We can change the conf.level to 0.99 instead of the default value 0.95

```{r}
t.test(bp_data$bp_before, bp_data$bp_after, paired = TRUE, conf.level = 0.99)
```

# t.test with formula
Sometimes we have data sets in the long format rather than wide, for example 

```{r}
bp_data_long <- pivot_longer(bp_data, cols = c(bp_before,bp_after), names_to = 'treatment', values_to = 'bp')
View(bp_data_long)
```

We can run t.test with formula
```{r}
paired_ttest_long <-  t.test(bp ~ treatment, data = bp_data_long, paired = TRUE)
```

Let's check if the results are the same

```{r}
paired_ttest$p.value == paired_ttest_long$p.value
paired_ttest$statistic == paired_ttest_long$statistic
```

Well the statistics are not the same, cause one is positive, the other is negative. Their ABSolute values should be the same

```{r}
abs(paired_ttest$statistic) == abs(paired_ttest_long$statistic)
```

# Visualisation 
have a look at the box-plot to confirm the t-test results

```{r}
ggplot(bp_data_long,aes(factor(treatment,levels = c('bp_before','bp_after')),bp)) +
  geom_boxplot()
```


```{r}
ggplot(bp_data_long,aes(factor(treatment,levels = c('bp_before','bp_after')),bp)) +
  geom_line(aes(group = patient_id))+
  geom_boxplot(width = 0.1)
```



