---
title: "Lab 14 review"
author: "Jian Zheng"
date: "`r Sys.Date()`"
output: html_document
---

Previous labs on my [github](https://github.com/jzheng23/INST314)

load packages (first install whichever you have not already installed)

```{r}
library(tidyverse)
library(effectsize)
library(pwr)
```

check the data

```{r}
#View(mpg)
glimpse(mpg)
```

# Proportion & bar-chart

## one proportion

Compare one proportion with the population parameter

prepare the data
```{r}
mpg$trans_auto = str_detect(mpg$trans,'auto*')
table(mpg$trans_auto)
```

test
```{r}
prop.test(157,234,p = 0.5)
```

viz
```{r}
ggplot(data = mpg, aes(x = 1, fill = trans_auto)) +
  geom_bar(position = 'fill') +
  geom_hline(yintercept = 0.5,color = 'red')
```

effect size & power
```{r}
pwr.p.test(ES.h(0.6709402,0.5),n = 234,sig.level = 0.05)
```


## two proportions
check the data
```{r}
table(mpg$year, mpg$trans_auto)
table(mpg$year)
```

test
```{r}
prop.test(c(74,83),c(117,117))
```


viz
```{r}
ggplot(data = mpg, aes(x = as.factor(year), fill = trans_auto)) +
  geom_bar(position = 'fill')
```

effect size & power
```{r}
pwr.p.test(ES.h(0.6324786, 0.7094017),n = 234,sig.level = 0.05)
```


# T-test & box-plot

## 1 sample t-test

t-test
```{r}
t.test(mpg$hwy, mu = 25)
```

viz
```{r}
ggplot(mpg, aes(x = 1, y = hwy))+
  geom_boxplot() +
  geom_hline (yintercept =  25,color = 'red')
```

effect size
```{r}
t_to_d(-4.0071, 233)
```

power
```{r}
pwr.t.test(234,-0.53,sig.level = 0.05,type = 'one.sample')
```

## Independent t-test
t test
```{r}
table(mpg$year)
t.test(hwy~year, data = mpg)
```

viz
```{r}
ggplot(mpg, aes(x = as.factor(year), y = hwy)) +
  geom_boxplot()
```

effect size
```{r}
t_to_d(-0.032864, 231.64)
```

power
```{r}
pwr.t.test(234,-4.32e-03,sig.level = 0.05, type = 'two.sample')
```


## Paired t-test
t-test
```{r}
t.test(mpg$cty, mpg$hwy, paired = TRUE)
```

reshape to long format to viz
```{r}
mpg_long <- mpg %>% 
  pivot_longer( cols = c(cty,hwy), names_to = 'road', values_to = 'mpg') %>%
  select('road','mpg')
#View(mpg_long)
```

viz
```{r}
ggplot(mpg_long, aes(x = road, y = mpg)) +
  geom_boxplot()
```

effect size
```{r}
t_to_d(-44.492, 233)
```

power
```{r}
pwr.t.test(234,d = -5.83,sig.level = 0.05,type = 'paired')
```


## ANOVA

ANOVA
```{r}
anova(lm(hwy ~ drv, data = mpg))
```

Pairwise comparison
```{r}
pairwise.t.test(mpg$hwy, mpg$drv, p.adjust.method = "bonferroni")
```

Box plot
```{r}
ggplot(mpg, aes(x = drv, y = hwy)) +
  geom_boxplot()
```

effect size
```{r}
F_to_f(f = 130.62,df = 2,df_error = 231)
```

power
```{r}
pwr.anova.test(k = 3,n = 234/3,f = 1.06)
```
# Chisq test & mosaic plot

check the data
```{r}
table(mpg$cyl, mpg$drv)
```

test
```{r}
chisq.test(table(mpg$cyl, mpg$drv))
```

viz
```{r}
#install.packages('ggmosaic')
library(ggmosaic)
ggplot(data = mpg) +
  geom_mosaic(aes(x = product(cyl), fill=drv))
```

effect size & power
```{r}
pwr.chisq.test(ES.w2(table(mpg$cyl, mpg$drv)/234),N = 234,df = 6)
```


# Regression & scatter plot

test
```{r}
result <- lm(hwy ~ cty, mpg)
summary(result)
```

viz
```{r}
ggplot(data = mpg, aes(x = cty, y = hwy))+
  geom_point() +
  geom_smooth(method = 'lm')
```


effect size
```{r}
eta2_to_f2(es = 0.9138)
```


power
```{r}
pwr.f2.test(u = 1,v = 232,f2 = 10.6)
```

# ggplot2 title

For more information, see [ggplot2 title : main, axis and legend titles](http://www.sthda.com/english/wiki/ggplot2-title-main-axis-and-legend-titles)

```{r}

ggplot(data = mpg, aes(x = cty, y = hwy))+
  geom_point() +
  geom_smooth(method = 'lm') +
  ggtitle('Highway MPG and city MPG') +
  xlab('City MPG') +
  ylab('Highway MPG') +
  theme_classic()

```
