---
title: "Lab_Week5_ggplot"
author: "Jian Zheng"
date: "2024-02-22"
output: html_document
---

# Introduction to ggplot2
ggplot2 is a powerful and a flexible R package for creating elegant data visualizations.

# Installing and Loading ggplot2

Installation: If you haven't installed ggplot2, you can do so by running install.packages("ggplot2"). 
You only need to install the package **once**, unless you uninstalled r. 

```{r}
#install.packages("ggplot2")
```

Loading: To use ggplot2, load it into your R session with library(ggplot2).
You need to load the package in each new session, e.g., in each new r or rmd file. 

```{r}
library(ggplot2)
```

# Basic Components of ggplot2

- Data: The dataset you want to visualize.
- Aesthetics (aes): Defines how variables in the data are mapped to visual properties (e.g., axes, colors, sizes).
- Geometries (geom_): The type of plot or layer you are adding (e.g., points, lines, bars).

# Creating a Basic Plot

## Start with ggplot() function

Define the dataset and aesthetics.
```{r}
#ggplot(data = your_dataset, aes(x = variable1, y = variable2))
ggplot(data = mtcars, aes(x = wt, y = mpg)) 
```

## Add geom_ layer

Specify the type of plot. E.g.,

Scatter plots: + geom_point()
```{r}
ggplot(data = mtcars, aes(x = wt, y = mpg)) +
  geom_point() +
  labs(title = "Scatter Plot of MPG vs Weight")
```

Bar charts: + geom_bar() 
```{r}
ggplot(mtcars, aes(x = factor(cyl))) +
  geom_bar() +
  labs(title = "Bar Chart of Cylinder Counts")
```

Histogram: + geom_histogram()
```{r}
ggplot(mtcars, aes(x = mpg)) +
  geom_histogram(binwidth = 3, fill = "green", alpha = 0.5, color = "black") +
  labs(title = "Histogram of MPG")
```

Box plot: + geom_boxplot()
```{r}
ggplot(mtcars, aes(x = factor(cyl), y = mpg)) +
  geom_boxplot() +
  labs(title = "Boxplot of MPG by Cylinder Count")
```

Density plots: + geom_density()
```{r}
ggplot(mtcars, aes(x = mpg)) +
  geom_density(fill = "green", alpha = 0.5) +
  labs(title = "Density Plot of MPG")
```

## Customize (optional) 

Add titles, labels, and modify themes.
```{r}
#+ ggtitle("Your Title Here")
#+ xlab("X-axis Label")
#+ ylab("Y-axis Label")
#+ theme_minimal()
ggplot(data = mtcars, aes(x = wt, y = mpg)) +
  geom_point() +
  ggtitle("Fuel Efficiency vs. Car Weight") +
  xlab("Weight (1000 lbs)") +
  ylab("Miles per Gallon") +
  theme_light()
```

# Sample code for ex 1

```{r}
# Generate data from a normal distribution
data <- data.frame(x = rnorm(10000))

# Create a density plot
ggplot(data, aes(x = x)) +
  geom_density(fill = "blue", alpha = 0.5) +
  labs(title = "Density Plot", x = "Value", y = "Density") +
  geom_vline(xintercept = 1, color = 'red') +
  geom_text(aes(x = 1.1, y = 0.1, label = "a")) +
  geom_vline(xintercept = -0.5, color = 'green')+
  geom_text(aes(x = -0.6, y = 0.1, label = "b"))
```

