---
title: "Lab_Week6_ggplotAdvanced"
author: "Jian Zheng"
date: "2024-02-27"
output: html_document
---

```{r}
library(ggplot2)
```


# Customizing Scales

Customize the scales of the scatter plot.
```{r}
ggplot(mtcars, aes(x = wt, y = mpg, color = as.factor(cyl))) +
  geom_point() +
  labs(title = "Custom Color MPG vs. Weight", color = "Cylinder Count") +
  scale_x_continuous(limits = c(0, 6), breaks = c(1, 3, 5)) +
  scale_y_continuous(limits = c(0, 35))
```

# Faceting for Multi-panel Plots

Create a multi-panel plot to compare mpg vs. wt across different levels of cyl using facet_wrap().

```{r}
p <- ggplot(data = mtcars, aes(x = wt, y = mpg)) +
  geom_point() +
  labs(title = "MPG vs. Weight, Faceted by Cylinder Count")
p + facet_wrap(vars(cyl))
```

Use facet_grid to specify the row and column.

```{r}
p + facet_grid(vars(am), vars(cyl))
```

# Combining Geom-Layers

Create a plot that combines scatter plots with a smooth linear model line.

```{r}
ggplot(data = mtcars, aes(x = wt, y = mpg)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(title = "MPG vs. Weight by Cylinder Count")
```

And a histogram with a density curve.

```{r message=FALSE}
ggplot(mtcars, aes(x = mpg)) +
  geom_histogram(aes(y=after_stat(density))) +
  geom_density() +
  labs(title = "MPG Distribution in mtcars Dataset",
       x = "Miles per Gallon",
       y = "Density") +
  theme_minimal()
```

# Statistical Transformations

Display a histogram of mpg with a custom bin width.

```{r message=FALSE}
ggplot(mtcars, aes(x = mpg)) +
  geom_histogram() +
  labs(title = "Histogram of MPG", x = "MPG", y = "Frequency")
```

Change Y axis to density.

```{r message=FALSE}
ggplot(mtcars, aes(x = mpg)) +
  geom_histogram(aes(y = after_stat(density))) +
  labs(title = "Histogram of MPG", x = "MPG", y = "Density")
```

See the difference?

# Arrange plots

Combine the two plots above into one

```{r message=FALSE}
library(cowplot)

p1 <- ggplot(mtcars, aes(x = mpg)) +
  geom_histogram(aes(y = after_stat(density))) +
  labs(title = "Histogram of MPG", x = "MPG", y = "Density")

p2 <- ggplot(mtcars, aes(x = mpg)) +
  geom_histogram() +
  labs(title = "Histogram of MPG", x = "MPG", y = "Frequency")

plot_col <- plot_grid(p1, p2, ncol = 1)

plot_col
```

Too flat (and wide). 

# Set size of plots

In r markdown, {r fig.height=5, fig.width=5}

```{r fig.height=5, fig.width=5}
plot_col
```

# Save the plot to a file

```{r}
ggsave("my_plot.png", width = 5, height = 5, units = "in") 
```
