---
title: "Week 4 R Markdown & pnorm"
author: "Jian Zheng"
date: "2024-02-15"
output: html_document
---
# Load packages

```{r}
library(dplyr)
library(ggplot2)
library(openintro)
```

# Norm

Normal distribution

## pnorm
The `pnorm()` function finds the *proportion* that is below a **certain value** on the Normal distribution. This, as with the other Normal distribution functions, have a default value of 

* 0 for the mean and 
* 1 for the standard deviation.

```{r}
pnorm(1)
```

```{r}
samplenorm <- rnorm(10000)
samplenorm_df <- data.frame(samplenorm)
samplenorm_df %>% ggplot(aes(samplenorm))+geom_density()+geom_vline(xintercept = 1)+geom_vline(xintercept = -1)
```
```{r}
pnorm(1) # below 1, on the left side of 1
1 - pnorm(1) # above 1, on the right side of 1
pnorm(1) - pnorm(-1) # between -1 and 1
```
## qnorm
We can use `qnorm()` to go the opposite way -- if we want to know what the cutoff is for a certain percentile.
```{r}
samplenorm_1 <- rnorm(10000, 10, 2)
```

```{r}
samplenorm_1_df <- data.frame(samplenorm_1)
samplenorm_1_df %>% ggplot(aes(samplenorm_1))+geom_density()
```
above 95% (or the top 5%)

```{r}
qnorm(.95, 10, 2)
```
```{r}
samplenorm_1_df %>% ggplot(aes(samplenorm_1))+geom_density()+geom_vline(xintercept =  13.289)
```

# Calculate the percentage between a range
`filer` and `nrow`

```{r}
data("email")
```

```{r}
email %>% ggplot(aes(x = line_breaks)) + geom_density()
```

Between 500 and 1000 lines
```{r}
email %>% filter((line_breaks < 1000) & (line_breaks > 500)) %>% nrow()
```

```{r}
nrow(email)
```
```{r}
493/3921
```

# R Markdown

1. Embedding R Code in Markdown Documents:
R code can be embedded in chunks ({r} ... ) to execute code and display results within the document.
Run a chunk or a line

2. Markdown Syntax:
Heading: Prefix the line with # for a heading. More #s indicate subheadings (e.g., ##, ###).
Bold: Wrap text with ** or __ (e.g., **bold**).
Italic: Wrap text with * or _ (e.g., *italic*).
List: Ordered and Unordered:
Unordered: Use - or * followed by a space.
Ordered: Use numbers followed by a period and space (e.g., 1. , 2. ).

3. YAML header
Configures document parameters like title, author, and output format.
Knitting Documents
Html, pdf, word

