---
title: "Lab_Week7_tTest"
author: "Jian Zheng"
date: "2024-03-06"
output: html_document
---

```{r message=FALSE}
library(ggplot2)
library(dplyr)
library(openintro)
```
# One-sample t-test

Is the mean of mpg different from 20?
```{r}
ggplot(mtcars, aes(x = mpg)) + geom_density() + geom_vline(xintercept = 20, color = "red") +
  labs(title = "Density of Miles per Gallon (mpg)", x = "Miles per Gallon (mpg)") 
```

```{r}
t.test(mtcars$mpg, mu = 20)
```

No, the mean of mpg is not different from 20.

# Two-sample t-test

Is the mpg of cars with automatic transmission different from that of cars with manual transmission?

```{r}
ggplot(mtcars, aes(x = factor(am), y = mpg, fill = factor(am))) +
  geom_boxplot() +
  labs(x = "Transmission (0 = Automatic, 1 = Manual)", y = "Miles per Gallon (mpg)", fill = "Transmission") 
```

## Using formula y ~ x

```{r}
t.test(mpg ~ am, data = mtcars)
```

## Comparing the two variables

```{r}
auto <- mtcars %>% filter(am == 0) %>% select(mpg)
manual <- mtcars %>% filter(am == 1) %>% select(mpg)

t.test(auto$mpg, manual$mpg)
```

Yes, cars with manual transmission have higher mpg than cars with automatifc transmission. 

# Paired t-test

Using the satgpa dataset from openintro
```{r}
data("satgpa")
```

```{r}
library(cowplot)
plot_v <- ggplot(satgpa, aes(y = sat_v)) + 
  geom_boxplot(fill = "skyblue") +
  ylim(c(min(c(satgpa$sat_v, satgpa$sat_m)), max(c(satgpa$sat_v, satgpa$sat_m)))) +
  labs(title = "SAT Verbal Scores", y = "SAT Verbal")+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())

plot_m <- ggplot(satgpa, aes(y = sat_m)) + 
  geom_boxplot(fill = "lightgreen") +
  ylim(c(min(c(satgpa$sat_v, satgpa$sat_m)), max(c(satgpa$sat_v, satgpa$sat_m)))) +
  labs(title = "SAT Math Scores", y = "SAT Math")+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())

plot_row <- plot_grid(plot_v, plot_m, ncol = 2)
plot_row
```
## A one-sample t-test on the difference

```{r}
sat_diff <- satgpa$sat_v - satgpa$sat_m
t.test(sat_diff)
```

## Comparing the two variables directly

```{r}
t.test(satgpa$sat_v, satgpa$sat_m, paired = TRUE)
```

